---
title: "AmNatHistoryModels"
author: "FD"
output: 
    html_document: 
          toc: true
          toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data

## Data sources

On 2020-07-20, I searched for all articles with `Publication Name = American Naturalist` on Web of Science Core collection. There were 11029 results. I manually exported all results in batches of 500 (export limit on WoS, when you also want to export the abstract), as `Tab-delimited (Mac)`. I collated the downloaded files, opened the merged file in Excel; I removed empty columns, renamed columns, as saved the file as `.csv`.

The source files are in <https://github.com/flodebarre/2020_AmNatHistory/tree/master/data/sourcesWOS/withAbstract>. 

## Load data 

```{r}
allArticles <- read.csv("../data/AmNat_allAbstracts.csv", stringsAsFactors = FALSE)
```

## Explore data

Number of articles loaded: `r nrow(allArticles)` items in the dataset.

```{r }
library(RColorBrewer)
cols <- brewer.pal(4, "Set2")
maincol <- cols[1]
```

Number of articles per year:
```{r}
nb.byY <- aggregate(allArticles$Authors, by = list(allArticles$PublicationYear), FUN = length)
par(las = 1)
plot(nb.byY$Group.1, nb.byY$x, 
     xlab = "Publication Year", ylab = "Number of items", 
     ylim = c(0, max(nb.byY$x)), axes = FALSE, 
     pch = 16, col = maincol)
axis(1, pos = 0)
axis(2, pos = 1900-1)
```

Articles with abstract:
```{r}
# Add information about number of characters in the abstract of each article
allArticles <- cbind(allArticles, AbsLength = nchar(allArticles$Abstract))

# Compute proportion of articles with non zero abstract length, by year
isAbs.byY <- aggregate((!is.na(allArticles$Abstract)), by = list(allArticles$PublicationYear), FUN = mean)

# Plot the result
par(las = 1)
plot(isAbs.byY$Group.1, isAbs.byY$x, 
     xlab = "Publication Year", ylab = "Proportion of items with non zero abstract length", 
     pch = 16, col = cols[2], 
     axes = FALSE)
axis(1, pos = 0)
axis(2, pos = 1900-1)
```

Subset of the data for which there are abstracts.
```{r}
absArticles <- allArticles[!is.na(allArticles$Abstract), ]
```

There are `r nrow(absArticles)` such items. 

# Search for words

Function to find a specific word in an abstract
```{r}
findWord <- function(word, line){
  grepl(pattern = word, x = absArticles[line, "Abstract"], ignore.case = TRUE)
}
```

Search for "model" in the abstracts
```{r}
v <- unlist(lapply(as.list(1:nrow(absArticles)), function(i) findWord("model", i)))

absArticles <- cbind(absArticles, wordModel = v)
```

Group by year and plot output
```{r}
model.byY <- aggregate(absArticles$wordModel, by = list(absArticles$PublicationYear), FUN = mean, na.rm = TRUE)

par(las = 1)
plot(range(model.byY$Group.1), range(model.byY$x), # Initialize the plot
     ylim = c(0, 1), 
     xlab = "Publication Year", ylab = "Proportion of abstracts containing the word `model`", 
     main = "``Model''",
     axes = FALSE,
     type = "n")
#rect(1900-1, 0, 2020, 1, col = gray(0.9), border = gray(0, 0))

#for(i in seq(0.1, 1, by = 0.1)){
#  lines(c(1900-1, 2020), rep(i, 2), col = gray(0.95), lwd = 1)
#}

axis(1, pos = 0)
axis(2, pos = 1900-1)

# Add the points
# (doing this so that the points are above the x axis...)
points(model.byY$Group.1, model.byY$x, pch = 16, col = maincol) 
```

